---
title: "Project 1:  predictive models"
author: 
- "Min He"
date: "July 2, 2020 (updated `r Sys.Date()`)"
output: 
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    code_folding: show
    toc_float:
      collapsed: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE)
library(readr)
library(tidyverse)
library(MASS)
library(caret)
library(randomForest)
```

## Introduction about the data

This dataset summarizes a heterogeneous set of features about articles published by Mashable in a period of two years. The goal is to predict the number of shares in social networks (popularity). [linked source] (https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity)

It includes the following columns (not all of them are described)

* shares: Number of shares in social media (this is the dependent variable)
*



## Read in the news popularity data

```{r}
setwd("/home/aaron/Documents/NCSU_MinHe/ST558/Project2")
pop_raw <- read.csv("OnlineNewsPopularity.csv", header = TRUE)
```

## Data preprocessing

* Combine the weekday flags into one variable `news_day`, e.g.: when `weekday_is_monday`=1, set the `news_day`='Monday'
* Combine the topic channels into one variable `data_channel`, e.g.: when `data_channel_is_lifestyle`=1, set the `data_channel`='lifestyle'
* Dividing the shares into two groups (< 1400 and ≥ 1400)

```{r}
# summary of the raw pop data
summary(pop_raw)
# From the summary, we can see there are outliers exist in valriables: n_tokens_content,n_unique_tokens,n_non_stop_words
# n_non_stop_unique_tokens, num_hrefs, num_self_hrefs, num_imgs, num_videos, kw_min_min, kw_max_min, kw_max_avg, kw_avg_avg
# self_reference_min_shares, self_reference_max_shares, self_reference_avg_sharess, shares

# Remove the non predictable variables

pop_raw$news_day <- rep("Sunday", nrow(pop_raw))
pop_raw$news_day[pop_raw$weekday_is_monday==1] <- "Monday"
pop_raw$news_day[pop_raw$weekday_is_tuesday==1] <- "Tuesday"
pop_raw$news_day[pop_raw$weekday_is_wednesday==1] <- "Wednesday"
pop_raw$news_day[pop_raw$weekday_is_thursday==1] <- "Thursday"
pop_raw$news_day[pop_raw$weekday_is_friday==1] <- "Friday"
pop_raw$news_day[pop_raw$weekday_is_saturday==1] <- "Saturday"
pop_raw$news_day <- factor(pop_raw$news_day) 
# convert the binary outcomes variables into factors

pop_raw$data_channel <- rep('', nrow(pop_raw))
pop_raw$data_channel[pop_raw$data_channel_is_lifestyle==1] <- "lifestyle"
pop_raw$data_channel[pop_raw$data_channel_is_entertainment==1] <- "entertainment"
pop_raw$data_channel[pop_raw$data_channel_is_bus==1] <- "business"
pop_raw$data_channel[pop_raw$data_channel_is_socmed==1] <- "socmed"
pop_raw$data_channel[pop_raw$data_channel_is_tech==1] <- "tech"
pop_raw$data_channel[pop_raw$data_channel_is_world==1] <- "world"
pop_raw$data_channel <- factor(pop_raw$data_channel) 

pop_raw$shares_flag <- pop_raw$shares>=1400
```

## Clearing data

* Remove the unused/non-predictive variables
* Remove outliers (e.g.: n_tokens_content<=0)
* Subset the records to the day defined by the argument

```{r}
pop_raw<-pop_raw %>% subset(select=-c(url, timedelta, is_weekend, weekday_is_monday, weekday_is_tuesday, weekday_is_wednesday, weekday_is_thursday, weekday_is_friday, weekday_is_saturday,weekday_is_sunday, data_channel_is_lifestyle, data_channel_is_entertainment, data_channel_is_bus, data_channel_is_socmed, data_channel_is_tech, data_channel_is_world, shares))

# filter out outliers
pop_raw_post <- pop_raw %>% filter( n_tokens_content > 0 & n_unique_tokens >= 0 & n_unique_tokens <=1.0 & n_non_stop_words>=0 & n_non_stop_words<=1.0 & n_non_stop_unique_tokens>=0 & n_non_stop_unique_tokens<=1.0)

pop_1_day <- pop_raw_post %>% filter(news_day == 'Monday') %>% subset(select=-c(news_day))
```

## Divide dataset into training v.s. testing datasets

* Set random seeds so that the results can be replicated
* Break the dataset into 70% for training v.s. 30% for testing

```{r}
# Sampling the dataset into 80% : training data and 20% : test data:
set.seed(100)
popNewsTrain <- sample(nrow(pop_1_day),as.integer(nrow(pop_1_day)*0.7))
train.news = pop_1_day[popNewsTrain,]
test.news = pop_1_day[-popNewsTrain,]
```

## Train the predictive model by using logistic regression

* Use both direction (forwards and backwards) stepwise methods to pick up the predict variables
* AIC is used as the criterion to pick the best combination of predict variables
* AIC estimates the relative amount of information lost by a given model: the less information a model loses, the higher the quality of that model. (excerpted from [linked wiki](https://en.wikipedia.org/wiki/Akaike_information_criterion))

```{r}
# Fit the full model 
# So try logistic regression model
full.model <- glm(shares_flag ~ ., data = train.news, family = "binomial")

# Stepwise regression model
step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
```

## Predict the value for test dataset with the trained logistic model

* If the predicted response is larger than 0.5, then it's shares>=1400
* Calculate/print the confusion matrix

```{r}
predict_value <- predict(step.model,test.news,type = "response") 
glm_confusionMatrix <- confusionMatrix(data = as.factor(predict_value>0.5), reference = as.factor(test.news$shares_flag))
glm_confusionMatrix
```

## Train the predictive model by using random forest

* Set number of trees to be 200


```{r}
bagFit <- randomForest(shares_flag ~ ., data = train.news, mtry = ncol(train.news) - 1,
                       ntree = 200, importance = TRUE)
```


## Predict the value for test dataset with the trained bagging model

* If the predicted response is larger than 0.5, then it's shares>=1400
* Calculate/print the confusion matrix

```{r}
bagPred <- predict(bagFit, newdata = dplyr::select(test.news, -shares_flag))

bag_confusionMatrix <- confusionMatrix(data = as.factor(bagPred>0.5), reference = as.factor(test.news$shares_flag))

```

## Conclusion
