---
title: "Project 2:  predictive models"
author: 
- "Min He"
date: "July 2, 2020 (updated `r Sys.Date()`)"
output: rmarkdown::github_document
---
  
```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE)
library(rmarkdown)
```


## Introduction about the data

This dataset summarizes a heterogeneous set of features about articles published by Mashable in a period of two years. The goal is to predict the number of shares in social networks (popularity). [source](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). In this project, I will try to use the logistic regression and bagging to classify whether the shares exceeds 1400 or not.

## Generate the report for all days automatically

```{r readme1}
#The unique days in a week
days <-  c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")

for(d in days){
  #create filenames
  output_file <- paste0( d, 'Analysis', ".md")
  #create a list for each analysis with the weekday parameter
  render("Project2.Rmd", output_file = output_file, params = list(weekday = d))
}

```


The analysis for [Monday is available here](MondayAnalysis.md)   
The analysis for [Tuesday is available here](TuesdayAnalysis.md)  
The analysis for [Wednesday is available here](WednesdayAnalysis.md)  
The analysis for [Thursday is available here](ThursdayAnalysis.md)  
The analysis for [Friday is available here](FridayAnalysis.md)  
The analysis for [Saturday is available here](SaturdayAnalysis.md)  
The analysis for [Sunday is available here](SundayAnalysis.md)  
